### 2020-09-23 tex/related-works.tex
Update related-works.tex
Added clarification w.r.t. how masking proceeds and a "defense" w.r.t. single trace attacks.

Power analysis attacks can be addressed using masking. Masking is one of the most widespread and better understood techniques to protect
against passive side-channel attacks. In its most basic form, a mask is drawn uniformly from random and added to the secret. The resulting masked value, which is effectively a one-time-pad, and the mask are jointly called \emph{shares}: if taken
singularly they are statistically independent from the secret, and they must be combined to obtain the secret back. Any operation that previously involved the secret has to be turned into an operation over its shares. As long as they are not combined, any leakage from them will be statistically independent of the secret too. In our context, we show how masking can easily be applied to FrodoKEM at a very low cost. We therefore argue the overhead that a protected implementation of Frodo in hardware incurs is minimal, hence making it a strong candidate when side-channel analysis are a concern. The reason behind this is that the only operation using the secret matrix $\mathbf{S}$ is the computation of the matrix $\mathbf{M}$ as $\mathbf{C} - \mathbf{B}'\mathbf{S}$ during decapsulation. When $\mathbf{S}$ is split in two (or more) shares using addition modulo $q$, the above multiplication by $\mathbf{B}'$ can be simply applied to both shares independently. Results are then subtracted by $\mathbf{C}$ one-by-one, so that computations never depend on both shares simultaneously.

>>>>

Power analysis attacks can be addressed using masking and hiding. Masking is one of the most widespread and better understood techniques to protect
against passive side-channel attacks. In its most basic form, a mask is drawn uniformly from random and added to the secret. The resulting masked value, which is effectively a one-time-pad, and the mask are jointly called \emph{shares}: if taken
singularly they are statistically independent from the secret, and they must be combined to obtain the secret back. Any operation that previously involved the secret has to be turned into an operation over its shares. As long as they are not combined, any leakage from them will be statistically independent of the secret too. In our context, we show how masking can easily be applied to FrodoKEM at a very low cost. We therefore argue the overhead that a protected implementation of Frodo in hardware incurs is minimal, hence making it a strong candidate when side-channel analysis are a concern. The reason behind this is that the only operation using the secret matrix $\mathbf{S}$ is the computation of the matrix $\mathbf{M}$ as $\mathbf{C} - \mathbf{B}'\mathbf{S}$ during decapsulation. When $\mathbf{S}$ is split in two (or more) shares using addition modulo $q$, the above multiplication by $\mathbf{B}'$ can be simply applied to both shares independently. Results are then subtracted by $\mathbf{C}$ one-by-one, so that computations never depend on both shares simultaneously. Concretely this implies that (in our two-share masking scheme that represents $\mathbf{S}$ as $(\mathbf{S_1},\mathbf{S_2})$) we compute $\mathbf{B}'\mathbf{S_1}$, followed by $\mathbf{M_1}=C-\mathbf{B}'\mathbf{S_1}$, and concluded by $\mathbf{M}=\mathbf{M_1}-\mathbf{B}'\mathbf{S_2}$. All terms in this computation are statistically independent of the secret.

Masking in itself is not sufficient to prevent all power/EM analysis attacks. For instance, single trace attacks (e.g. such as horizontal attacks or template attacks), can, if a devices signal to noise ratio is favourable, extract information from even a single side channel observation. If this is possible, then in the worst case all shares can be extracted and therefore masking becomes ineffective. Circumventing such attacks always require to make the signal to noise ratio less favourable, for instance, by ensuring that idle hardware contributes to the power or EM consumption. If such control on the hardware level is not possible to hiding countermeasures such as re-ordering of (basic) operations or dummy operations can be used. In our work we are interested in a design space exploration: this means our focus is on the impact of countermeasures on the architecture of our basic design. Architectural changes are more likely incurred by dealing with masking then by turning on/off components or by hiding countermeasures.
